---
title: "Explore Table 502.20 (Annual earnings across levels of education)"
author: "Winnie Wing-Yee Tse"
date: "4/3/2022"
output: pdf_document
---

```{r}
library(tidyverse)
```


Median annual earnings, number, and percentage of full-time year-round workers age 25 and over, by highest level of educational attainment and sex: Selected years, 1990 through 2019

0. Total
1. Less than 9th grade
2. Some high school, no completion
3. High school completion (includes equivalency)
4. Some college, no degree
5. Associate's degree
6. Bachelor's or higher (total)
7. Bachelor's degree
8. Master's degree
9. Professional degree
10. Doctor's degree


```{r data}
names_col <- c("year", 
               paste0(c("mean_", "se_"), rep(0:10, each = 2)))
# Current dollors
cd_m <- readxl::read_excel("../../data/tabn502.20.xls", 
                           range = "A10:W31", col_names = names_col)
cd_f <- readxl::read_excel("../../data/tabn502.20.xls", 
                           range = "A33:W54", col_names = names_col)

# Constant 2019 dollars
cd2019_m <- readxl::read_excel("../../data/tabn502.20.xls", 
                               range = "A57:W78", col_names = names_col)
cd2019_f <- readxl::read_excel("../../data/tabn502.20.xls", 
                               range = "A80:W101", col_names = names_col)

# Number of persons with earnings who worked full time, year round (in thousands)
nft_m <- readxl::read_excel("../../data/tabn502.20.xls", 
                            range = "A104:W125", col_names = names_col)
nft_f <- readxl::read_excel("../../data/tabn502.20.xls", 
                            range = "A127:W148", col_names = names_col)

# Percent of persons with earnings who worked full time, year round
perft_m <- readxl::read_excel("../../data/tabn502.20.xls", 
                              range = "A151:W170", col_names = names_col)
perft_f <- readxl::read_excel("../../data/tabn502.20.xls", 
                              range = "A172:W191", col_names = names_col)
```

```{r}
longtab <- function(widetab) {
  widetab %>%
    # force non-numeric symbols to NA
    mutate_at(vars(year:se_10), as.numeric) %>%
    pivot_longer(
      mean_0:se_10, 
      names_to = c("type", "education"), 
      values_to = "est", 
      names_pattern = "(mean|se)_(.*)"
    ) %>%
    suppressWarnings()
}
```

```{r}
cd <- data.frame(gender = rep(c("male", "female"), each = nrow(cd_m)), 
                 rbind(cd_m, cd_f)) %>%
  longtab()
cd2019 <- data.frame(gender = rep(c("male", "female"), each = nrow(cd_m)), 
                 rbind(nft_m, nft_f)) %>%
  longtab()
nft_m <- data.frame(gender = rep(c("male", "female"), each = nrow(cd_m)), 
                 rbind(nft_m, nft_f)) %>%
  longtab()
```



